---
title: "Activity Monitoring"
author: "Maisoon"
date: "November 19, 2020"
output:
  html_document: default

---
```{r}
knitr::opts_chunk$set(echo = TRUE,fig.path="figures/")
```
#Introduction:

#### We want to record and analyse the steps of an individual per day for two months.


##1.Loading and preprocessing the data
```{r}
URL="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(URL,destfile="./stepsdata")
unzip("stepsdata")
activity=read.csv("activity.csv")
head(activity)
```


##2.What is mean total number of steps taken per day?

### Total number of steps taken per day
```{r}
activity$date=as.Date(activity$date)  ##change class of date to "Date"
stepsperday=tapply(activity$steps, activity$date, sum)
head(stepsperday)
```

### Histogram of the total number of steps taken each day
```{r}
hist(stepsperday,breaks=30,label=FALSE,main="Daily Steps",xlab="Steps",col="red")
```

### Mean and Median of the total number of steps taken per day after removing NA
mean:
```{r}
mean(stepsperday,na.rm=TRUE)
```
median:
```{r}
median(stepsperday,na.rm=TRUE)
```


##3. What is the average daily activity pattern?

### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
AverageActivity=aggregate(steps~interval,activity,mean,na.rm=TRUE)
library(ggplot2)
ggplot(AverageActivity,aes(interval,steps))+geom_line(col="red")+ggtitle("Average Activity Pattern")
```

### 5-minute interval containing the maximum number of steps 
```{r}
AverageActivity[AverageActivity$steps==max(AverageActivity$steps),1]
```


##4. Imputing missing values

### Total number of missing values in the dataset 
```{r}
sum(is.na(activity$steps))
```

### Filling in all of the missing values in the dataset by using mean of the 5-min interval
```{r}
merge=merge(activity,AverageActivity,by="interval")
for(i in 1:nrow(activity)){
    if(is.na(merge[i,"steps.x"])){
    merge[i,"steps.x"]=merge[i,"steps.y"]
    }
}
head(merge)
```

### Creating new dataset that is equal to the original dataset but with the missing data filled in
```{r}
ActivityNoNA=data.frame(steps=merge$steps.x,date=merge$date,interval=merge$interval)
head(ActivityNoNA)
```

### Histogram of the total number of steps taken each day
```{r}
ActivityNoNA$date=as.Date(ActivityNoNA$date)
NoNAstepsperday=with(ActivityNoNA,tapply(steps,date,sum))
hist(NoNAstepsperday,label=FALSE,breaks=30,xlab="Steps",col="red",main="Daily Steps")
```

#### The mean and median total number of steps taken per day after replacing NA
mean:
```{r}
mean(NoNAstepsperday)
```
median:
```{r}
median(NoNAstepsperday)
```
Conclusion: 

-The mean of the steps per day stayed the same after the replacements of NA

-The median increased by about 0.01% after the replacements of NA


##5.Are there differences in activity patterns between weekdays and weekends?

### Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
ActivityNoNA$day=rep(NA,nrow(ActivityNoNA))
for(i in 1:nrow(ActivityNoNA)){
                 if(weekdays(ActivityNoNA$date[i])=="Satuday"|
                    weekdays(ActivityNoNA$date[i])=="Sunday"){
                    ActivityNoNA$day[i]="Weekend"
                  }else{
                    ActivityNoNA$day[i]="Weekday"
                  }
}
ActivityNoNA$day=factor(ActivityNoNA$day,labels = c("Weekday","Weekend"))
head(ActivityNoNA)
```

### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
meanperday=aggregate(steps~interval+day,ActivityNoNA,mean)
ggplot(meanperday,aes(interval,steps,color=day))+geom_line()+facet_wrap(.~day,ncol=1,nrow=2)+ggtitle("Average of Steps across Weekdays and Weekends")
```



